Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-11 article/article.tex
Commit 2024-06-11 09:11

According to this study, in the period 1997 -- 2005, the mean AOD at
\(340\,\text{nm}\) \(0.403\) with a trend of \(-3.8\pm0.93\,\%\) per
year, corresponding to a change of \(0.0153\) per year. Using an
Ångström coefficient \(\alpha = 1.6\), this translates to a change in
the Ångström coefficient \(\beta=0.00272\) per year (or \(\beta=0.084\)
in 1997 and \(\beta=0.059\) in 2005). Simulations with uvspec for the
above Ångström coefficients, with WC of \(15.7\,\text{mm}\) and
\(15.4\,\text{mm}\) for 1997 and 2005 respectively and for a SZA of
\(55^\circ\) reveal trend of \(+0.21\,\%\) per year in
\(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was chosen as
representative of all days in the year in order to get a rough estimate
of the annually averaged change in clear sky irradiance. For the period
2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET
in a similar simulation scheme to calculate the monthly mean of clear
sky irradiance, and finally the trend of \(+0.14\,\%\) per year. We
applied these two long term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

First, we evaluated the performance of the modelled
\(\text{GHI}_\text{ref}\) in relation to the measured GHI for different
levels of atmospheric clearness, by using in the RTM the monthly
climatological AOD and WC, less their respective standard deviations.
These values represent typical atmosphere in Thessaloniki, with lower
than average load of aerosols and humidity, which are the main factor
that attenuate the GHI, excluding clouds. To create the upper limit of
CE identification, we used a value relative to clear sky irradiance
(\(4\,\%\)) with an additional constant offset of
\(15\,\text{W}/\text{m}^2\), as described in
Equation\nobreakspace\ref{eq:CE4}). \begin{equation}
\text{CE} : \text{E}_\text{i} > 15 + 1.04 \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation} where: \(\text{E}\) the measured irradiance,
\(\text{E}_\text{CSm}\) the selected modelled clear sky irradiance, and
\(i\) each of the one-minute observation. This is the criterion of our
CE identification.

In the process of defining the above details, we implemented an
empirical method, by manual inspection of the CE identification results,
on selected days from the whole dataset. We used seven sets of days,
with characteristics relevant to the efficiency of the identification
threshold. These sets were random groups of about 20 to 30 days with the
following characteristics: (a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance, (c) absence of clouds (by
implementing a clear sky identification algorithm as discussed in
\citet{Natsis2023}), (d) absence of clouds and absence of EC events, (e)
with at least \(60\,\%\) of the day length without clouds and presence
of EC events, (h) random selected days and (i) some manual selected days
that were included during the manual inspection process. Where it was
needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

The definition of the CE events with this method, has a level of
subjectivity. The actual clear sky irradiance, is not known, and can
only be approximated. Although, because our main focus is to be able to
identify the OIR created by the clouds, we expect that we were able to
detect all the major events of OIR due to CE. Where, some CE events,
with very low OIR, may be not detected, these are few and their
magnitude will not effect our results.

Another aspect of the CE events, that is often reported in the relative
bibliography, are cases of extreme cloud enhancement (ECE). These are
cases of CE, where the measured intensity of the irradiance, exceeds the
equivalent TSI on the top of the atmosphere, and detected according to
the Equation\nobreakspace{}\ref{eq:ECE}. \begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation} where: \(\theta\) the solar zenith angle, \(E_{\odot}\)
the solar constant, \(r\) the Sun -- Earth distance, and \(i\) each of
the one-minute observation.

>>>>

According to this study, in the period 1997 -- 2005 the mean AOD at
\(340\,\text{nm}\) \(0.403\) with a trend of \(-3.8\pm0.93\,\%\) per
year, corresponding to a change of \(0.0153\) per year. Using an
Ångström coefficient \(\alpha = 1.6\), this translates to a change in
the Ångström coefficient \(\beta=0.00272\) per year (or \(\beta=0.084\)
in 1997 and \(\beta=0.059\) in 2005). Simulations with uvspec for the
above Ångström coefficients, with WC of \(15.7\,\text{mm}\) and
\(15.4\,\text{mm}\) for 1997 and 2005 respectively and for a SZA of
\(55^\circ\) reveal trend of \(+0.21\,\%\) per year in
\(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was chosen as
representative of all days in the year in order to get a rough estimate
of the annually averaged change in clear sky irradiance. For the period
2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET
in a similar simulation scheme to calculate the monthly mean of clear
sky irradiance, and finally the trend of \(+0.14\,\%\) per year. We
applied these two long term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

First, we evaluated the performance of the modelled
\(\text{GHI}_\text{ref}\) in relation to the measured GHI for different
levels of atmospheric clearness, by using in the RTM the monthly
climatological AOD and WC, less their respective standard deviations.
These values represent typical atmosphere in Thessaloniki, with lower
than average load of aerosols and humidity, which are the main factor
that attenuate the GHI, excluding clouds. With this approach the
simulated \(\text{GHI}_\text{ref}\) should be generally greater than the
measured GHI when aerosols are more abundant. To compensate for this, we
increased the \(\text{GHI}_\text{ref}\) by (\(4\,\%\)) with an
additional constant offset of \(15\,\text{W}/\text{m}^2\), as described
in Equation\nobreakspace\ref{eq:CE4}. \begin{equation}
\text{CE} : \text{E}_\text{i} > 15 + 1.04 \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation} where: \(\text{E}\) the measured irradiance,
\(\text{E}_\text{CSm}\) the selected modelled clear sky irradiance, and
\(i\) each of the one-minute observation. This is the criterion of our
CE identification.

These values were determined through the implementation of an empirical
method with manual inspection of the CE identification results on
selected days of the whole dataset. We used seven sets of days with
characteristics relevant to the efficiency of the identification
threshold. These sets were random groups of about 20 -- 30 days with the
following characteristics: (a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance, (c) absence of clouds (by
implementing a clear sky identification algorithm as discussed in
\citet{Natsis2023}), (d) absence of clouds and absence of EC events, (e)
with at least \(60\,\%\) of the day length without clouds and presence
of EC events, (h) randomly selected days and (i) manually selected days.
Where needed in some of the edge cases, we also used images from a
sky-camera to further aid the decision of the manual inspection.

The definition of the CE events with this method has a degree of
subjectivity, since the actual clear sky irradiance, is not known, and
can only be approximated. However, this method is capable in detecting
all major CE events. Where, some CE events, with very low OIR may be not
detected, these are few with small over-irradiance and it is unlikely
that will effect significantly our results.

A sub-category of the CE events that is often discussed in the relevant
bibliography, are cases of extreme cloud enhancement (ECE). These are
cases of CE, where the measured intensity of the irradiance, exceeds the
equivalent TSI on the top of the atmosphere, and detected according to
the Equation\nobreakspace{}\ref{eq:ECE}. \begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation} where: \(\theta\) the solar zenith angle, \(E_{\odot}\)
the solar constant, \(r\) the Sun -- Earth distance, and \(i\) each of
the one-minute observation.
